---
title: 'Final Project'
author: "Swarali Korgaonkar & Akanksha Bhusari"
date: "Due April 30, 2019"
output: html_notebook
---

# Front matter

```{r echo=TRUE, message=FALSE}
# always clean up R environment
rm(list = ls())
# load all packages here
library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(caret)

require(rpart)
require(partykit)

# user-defined functions here (if any)
clean_data <-
  function(data){
    country <-
      data %>%
        select(1, 3)
    country <-
      unique(country)
    
    AVG_data <-
      data %>%
        group_by(Name) %>%
        summarize(world_rank=mean(world_rank), yr2012=sum(yr2012, na.rm = T), yr2013=sum(yr2013, na.rm = T), yr2014=sum(yr2014, na.rm = T), yr2015=sum(yr2015, na.rm = T), quality_of_education=mean(quality_of_education), alumni_employment=mean(alumni_employment), quality_of_faculty=mean(quality_of_faculty), publications=mean(publications), influence=mean(influence), citations=mean(citations), broad_impact=mean(broad_impact), patents=mean(patents), score=mean(score), Applicants.total=mean(Applicants.total), Admissions.total=mean(Admissions.total), TotalEnrollment=mean(TotalEnrollment), AcceptanceRate=mean(AcceptanceRate), TribalCollege=mean(TribalCollege), AreaType.UCL=mean(AreaType.UCL), Enrollment=mean(Enrollment), AmericanIndianAlaska.Pct=mean(AmericanIndianAlaska.Pct), Asian.Pct=mean(Asian.Pct), Black.Pct=mean(Black.Pct), HispanicLatino.Pct=mean(HispanicLatino.Pct), HawaiianPacificIslander.Pct=mean(HawaiianPacificIslander.Pct), White.Pct=mean(White.Pct), TwoMoreRace.Pct=mean(TwoMoreRace.Pct), UnknownEthnicity.Pct=mean(UnknownEthnicity.Pct), NonResidentAlien.Pct=mean(NonResidentAlien.Pct), AsianHawaiianPacificIslander.Pct=mean(AsianHawaiianPacificIslander.Pct), Women.Pct=mean(Women.Pct), FirstUndergradForeign.Num=mean(FirstUndergradForeign.Num), FirstUndergradForeign.Pct=mean(FirstUndergradForeign.Pct))
    
    TOTALdata <-
      merge(x = AVG_data, y = country, by = "Name", all = TRUE)
    
    return(TOTALdata)
    
  }


# load data
```

# Download datasets and Data Preprocessing 

```{r}
# CollegeRankings <-
#   read.csv(file="/Users/Swarali/Desktop/STAT 380/final-project/FinalData/world-university-rankings/cwurData.csv", header=TRUE, sep=",")
CollegeRankings <-
  read.csv(file="C:/Users/Akanksha Bhusari/Documents/STAT380/final-project/cwurData.csv", header=TRUE, sep=",")

names(CollegeRankings)[names(CollegeRankings) == 'institution'] <- 'Name'

CollegeRankings

# CollegeAdmissions <- 
#   read.csv(file="/Users/Swarali/Desktop/STAT 380/final-project/FinalData/Data-Table1.csv", header=TRUE, sep=",")
CollegeAdmissions <- 
  read.csv(file="C:/Users/Akanksha Bhusari/Documents/STAT380/final-project/Data-Table 1.csv", header=TRUE, sep=",")

#Subset data to only include columns needed
CollegeAdmissions_tidy <- 
  CollegeAdmissions %>%
  select(1:4, 28, 42:44, 46, 53:63, 90:91)

#Change column names
setnames(CollegeAdmissions_tidy, 
         old=c("Enrolled.total", "Percent.admitted...total", "Historically.Black.College.or.University", "Tribal.college", "Degree.of.urbanization..Urban.centric.locale.", "Total..enrollment", "Percent.of.total.enrollment.that.are.American.Indian.or.Alaska.Native", "Percent.of.total.enrollment.that.are.Asian", "Percent.of.total.enrollment.that.are.Black.or.African.American", "Percent.of.total.enrollment.that.are.Hispanic.Latino", "Percent.of.total.enrollment.that.are.Native.Hawaiian.or.Other.Pacific.Islander", "Percent.of.total.enrollment.that.are.White", "Percent.of.total.enrollment.that.are.two.or.more.races", "Percent.of.total.enrollment.that.are.Race.ethnicity.unknown", "Percent.of.total.enrollment.that.are.Nonresident.Alien", "Percent.of.total.enrollment.that.are.Asian.Native.Hawaiian.Pacific.Islander", "Percent.of.total.enrollment.that.are.women", "Number.of.first.time.undergraduates...foreign.countries", "Percent.of.first.time.undergraduates...foreign.countries"), 
         new=c("TotalEnrollment", "AcceptanceRate", "HBCU", "TribalCollege", "AreaType.UCL", "Enrollment", "AmericanIndianAlaska.Pct", "Asian.Pct", "Black.Pct", "HispanicLatino.Pct", "HawaiianPacificIslander.Pct", "White.Pct", "TwoMoreRace.Pct", "UnknownEthnicity.Pct", "NonResidentAlien.Pct", "AsianHawaiianPacificIslander.Pct", "Women.Pct", "FirstUndergradForeign.Num", "FirstUndergradForeign.Pct")) 

CollegeAdmissions_tidy
```

# Join datasets (Left Join)

Joins (at least) two different data sources for analysis
Report includes meaningful contributions from (at least) two different data sources that are JOINED for the analysis

```{r}
total_data<-
  merge(x=CollegeRankings,y=CollegeAdmissions_tidy,by="Name",all.x=TRUE)
total_data
```
The reason to why there are multiple listings of the universities is because the data includes statistics from multiple years and the data per university changes yearly.

# Data Wrangling

Clear demonstration of proficiency should include proper use of 5 out of the following 6 topics from class: (+) general purpose data wrangling--e.g., filter, select mutate, summarise, arrange, group_by, etc. (+) spread & gather to stack/unstack variables (+) regular expressions (+) user-defined functions (+) loops and control flow (+) vectorized functions like the `apply` family, `purrr::map` or `dplyr::do`

## filter, select, mutate, summarise, arrange, group_by, etc

Within this section of the code, we decided to filter the universities that were ranked within the top 100 nationally in order to later determine how much impact diversity plays in a school's success. 

## spread & gather to stack/unstack variables

```{r}
total_data <-
  spread(total_data, year, national_rank, fill = NA)

total_data
```
```{r}
# rename columns 
names(total_data)[37]<-
  "yr2015"
names(total_data)[36]<-
  "yr2014"
names(total_data)[35]<-
  "yr2013"
names(total_data)[34]<-
  "yr2012"

total_data
```

## regular expressions



## user-defined functions

```{r}

FINALdata <-
  clean_data(total_data)
FINALdata
```


## loops and control flow
```{r}
top_schools <-
  filter(FINALdata, world_rank < 100)

top_schools
```

```{r}
countries <-
  top_schools %>%
    select(36)
countries <-
  unique(countries)

countries

```

```{r}
countriesList <-
  c(countries[1])
countriesList
```

```{r}
countries
```



## vectorized functions like the `apply` family, `purrr::map` or `dplyr::do`


```{r}

variables <-
  top_schools %>%
    select(2, 7, 9, 10, 11, 14)

for (x in variables){
  V <-
    tapply(x, top_schools$country, FUN=mean)

  V <-
    as.data.frame(V)

  V <-
    cbind(country = rownames(V), V)

  countries <-
    merge(x = countries, y = V, by = "country", all.x = TRUE)
  
}

names(countries)[2]<-
  "AvgWorldRank"
names(countries)[3]<-
  "AvgQEdu"
names(countries)[4]<-
  "AvgQFaculty"
names(countries)[5]<-
  "AvgPublications"
names(countries)[6]<-
  "AvgInfluence"
names(countries)[7]<-
  "AvgPatents"

countries
```


# Data Analysis
Clear demonstration of proficiency should include proper use of 3 out of the following 5 topics from class: (+) statistical modeling/supervised learning (+) unsupervised learning (+) user-defined simulations (+) analysis of text data (+) R tools for "big data"

## statistical modeling/supervised learning

```{r}
pairs(countries)
countries
```

```{r}

mod_tree <- 
  rpart(country ~ AvgQEdu + AvgQFaculty + AvgPublications + AvgInfluence + AvgPatents, data = countries)

tree_plot <-
  plot(as.party(mod_tree))

tree_plot

```

```{r}

printcp(mod_tree)
plotcp(mod_tree)

```

## unsupervised learning
```{r}
result <- 
  countries[-1]
row.names(result) <- 
  countries$country
result
```

```{r}
#countries$country = as.numeric(as.factor(countries$country))

countries_pca <- result %>%
  prcomp(scale = TRUE)  


(-1) * countries_pca$rotation[, 1:2] %>% 
  round(2)

```
```{r}
countries_pca$x %>%
  as.data.frame() %>%  # `ggplot2` expects a data frame object
  ggplot(aes(x = PC1, y = PC2)) + 
  geom_point(aes(color = countries$country), size = 3) + 
  xlab("Best Vector from PCA") + 
  ylab("Second Best Vector from PCA") + 
  ggtitle("Two-dimensional representation of Average Country Education values (colored by country)") 
```

```{r}
countries_std <-
  scale(result) %>%
  as.data.frame()

countries_dist <- 
  dist(countries_std)

countries_dendo <-
  countries_dist %>%
    hclust(method = "complete")

countries_dendo %>%
  plot(cex = 0.9, labels = countries$country, lwd = 2,
       main = "Countries Dendogram with Complete Linkage")

```


## user-defined simulations

```{r}

```


## analysis of text data

```{r}

```


## R tools for "big data"

```{r}

```


# Data Visualization

Clear demonstration of proficiency should include at least FOUR distinct and useful data visualizations that are relevant to purpose of the analysis. Include at least one effective visualization with layered data from distinct data structures--e.g., a `geom` that plots data from a secondary data frame AND another must effectively display many--3 or more--variables. All plots must demonstrate good plotting practices.

```{r}
top_schools %>%
  ggplot(aes(x = quality_of_faculty, y = quality_of_education)) + 
  geom_point(aes(color = country), size = 3) + 
  xlab("Quality of Faculty") + 
  ylab("Quality of Education") + 
  ggtitle("How Quality of Faculty Affects Quality of Edu by Country for the world's Top 100 Universities")

reg <- 
  lm(quality_of_education~quality_of_faculty,data=top_schools) 
with(top_schools,plot(quality_of_faculty, quality_of_education))
	abline(reg)
```

```{r}
top_USA_schools <-
  filter(top_schools, country == "USA")
top_USA_schools

```















